### 2021-01-27 introduction.tex
Minor layout changes.

In this paper we make the following contributions:
\begin{enumerate}
\item
\textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}) as well as other examples.

 \item
 \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate}
	\item our version does not have the pseudo-type $S^\sharp$;
	\item our structural congruence is type preserving.
	\end{enumerate}
\item
\textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi. %We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).

\item
\textbf{Connection to linear logic}. The connection of PGV to logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).  %We showcase it with the encoding of cyclic scheduler to PGV (\cref{sec:pgv-example}).
\end{enumerate}
\end{document}

>>>>

In this paper we make the following contributions:
\begin{enumerate}
\item \textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
\item \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate*}
	\item our version does not have the pseudo-type $S^\sharp$;
	\item our structural congruence is type preserving.
	\end{enumerate*}
\item \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi.
\item \textbf{Connection to linear logic}. The connection of PGV to logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
\end{enumerate}
\end{document}

### 2021-01-27 priority-gv.tex
Minor layout changes.

In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when does it start communicating, and when does it finish. The upper bound is written on the sequent, whereas the lower bound is approximated from the typing environment. Typing rules for sequential constructs enforce sequentially, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
\begin{mathpar}
  \inferrule*{
    \inferrule*{
      \inferrule*{
      }{
        \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{y'}{\tyrecv{\tyunit}{\tyendr}}
      }
    }{
      \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{\recv\;{y'}}{\typrod{\tyunit}{\tyendr}}
    }
    \\
    \inferrule*{
    }{
      \tseq{}{\send\;{\pair{\unit}{x}}}{}
    }
  }{
    \tseq{}{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}}{}
  }

  {\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
\end{mathpar}
Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them. For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
\begin{mathpar}
  \inferrule*{
    \inferrule*{
      \inferrule*{}{\tm{\send}}
      \\
      \inferrule*{
        \inferrule*{
        }{
          \tseq[\pbot]{\tmty{x}{A}}{x}{A}
        }
        \\
        \inferrule*{
        }{
          \tseq[\pbot]
          {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
          {y}{\tysend[\cs{o}]{A}{S}}
        }
      }{
        \tseq[\pbot]
        {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
        {\pair{x}{y}}{\typrod{A}{\tysend[\cs{o}]{A}{S}}}
      }
    }{
      \tseq[\cs{o}]
      {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
      {\send\;\pair{x}{y}}{S}
    }
  }{
    \tseq[\pbot]
    {\tmty{y}{\tysend[\cs{o}]{A}{S}}}
    {\lambda{x}.\send\;\pair{x}{y}}{\tylolli[\cs{o},\cs{o}]{A}{S}}
  }
\end{mathpar}
The approximation of the lower bound is not entirely accurate, \eg, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ has lower bound $\ptop$, not $\cs{o}$, but as we only ever under-approximate, it is safe.

>>>>

In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when does it start communicating, and when does it finish. The upper bound is written on the sequent, whereas the lower bound is approximated from the typing environment. Typing rules for sequential constructs enforce sequentially, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts. The typing rule for $\tm{\new}$ ensures that both endpoints of a channel share the same priorities. Together, these two constraints guarantee deadlock freedom.
For instance, the program in \cref{ex:deadlock} is deadlocked. Crucially, it composes the terms below in parallel. While each of these terms itself is well-typed, they impose opposite conditions on the priorities, so their composition is ill-typed. (We omit the priorities on $\ty{\tyends}$ and $\ty{\tyendr}$.)
\begin{mathpar}
  \small
  \inferrule*{
    \tseq[\cs{o'}]
    {\tmty{y'}{\tyrecv[\cs{o'}]{\tyunit}{\tyendr}}}
    {\recv\;{y'}}
    {\typrod{\tyunit}{\tyendr}}
    \\
    \tseq[\cs{p}]
    {\tmty{x}{\tysend[\cs{o}]{\tyunit}{\tyends}},\tmty{y'}{\tyendr}}
    {\letbind{x}{\send\;{\pair{\unit}{x}}}{\dots}}
    {\tyunit}
    \and
    \cs{o'}<\cs{o}
  }{%
    \tseq[\cs{p}]
    {%
      \tmty{x}{\tysend[\cs{o}]{\tyunit}{\tyends}},
      \tmty{y'}{\tyrecv[\cs{o'}]{\tyunit}{\tyendr}}
    }
    {\letpair{\unit}{y'}{\recv\;{y'}}{\letbind{x}{\send\;{\pair{\unit}{x}}}{\dots}}}
    {\tyunit}
  }

  \inferrule*{
    \tseq[\cs{o}]
    {\tmty{x'}{\tyrecv[\cs{o}]{\tyunit}{\tyendr}}}
    {\recv\;{x'}}
    {\typrod{\tyunit}{\tyendr}}
    \\
    \tseq[\cs{q}]
    {\tmty{y}{\tysend[\cs{o'}]{\tyunit}{\tyends}},\tmty{x'}{\tyendr}}
    {\letbind{y}{\send\;{\pair{\unit}{y}}}{\dots}}
    {\tyunit}
    \and
    \cs{o}<\cs{o'}
  }{%
    \tseq[\cs{q}]
    {%
      \tmty{y}{\tysend[\cs{o'}]{\tyunit}{\tyends}},
      \tmty{x'}{\tyrecv[\cs{o}]{\tyunit}{\tyendr}}
    }
    {\letpair{\unit}{x'}{\recv\;{x'}}{\letbind{y}{\send\;{\pair{\unit}{y}}}{\dots}}}
    {\tyunit}
  }
\end{mathpar}
Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them. For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
\begin{mathpar}
  \small
  \inferrule*{
    \inferrule*{
      \inferrule*{
      }{
        \tmty{\send}{\tylolli[\ptop,\cs{o}]{\typrod{A}{\tysend[\cs{o}]{A}{S}}}{S}}
      }
      \and
      \inferrule*{
        \inferrule*{
        }{
          \tseq[\pbot]{\tmty{x}{A}}{x}{A}
        }
        \and
        \inferrule*{
        }{
          \tseq[\pbot]
          {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
          {y}{\tysend[\cs{o}]{A}{S}}
        }
      }{
        \tseq[\pbot]
        {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
        {\pair{x}{y}}{\typrod{A}{\tysend[\cs{o}]{A}{S}}}
      }
    }{
      \tseq[\cs{o}]
      {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
      {\send\;\pair{x}{y}}{S}
    }
  }{
    \tseq[\pbot]
    {\tmty{y}{\tysend[\cs{o}]{A}{S}}}
    {\lambda{x}.\send\;\pair{x}{y}}{\tylolli[\cs{o},\cs{o}]{A}{S}}
  }
\end{mathpar}
The approximation of the lower bound is not entirely accurate, \eg, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ has lower bound $\ptop$, not $\cs{o}$, but as we only ever under-approximate, it is safe.
